Principles for Open Data Curation: A Case Study with the New York City 311 Service Request Data

### 2024-10-23 manuscript/manuscript.tex
Update to date aggregation and review of manuscript
Made changes to date aggregations (monthly, daily, hourly, etc.) to improve performance. Altered two images in manuscript. Continued editing.

The Mostly Empty category includes such fields as
\texttt{taxi\_company due\_date},
\texttt{pickup\_location}, and \texttt{landmark}.
The Partially Empty includes such fields as
\texttt{location\_type}, \texttt{borough}, 
and \texttt{cross\_street\_1 \& \_2}. And the Few/None Empty includes
\texttt{created\_date}, \texttt{complaint\_type},
\texttt{agency}, and \texttt{status}. In some analysis efforts, it may be
prudent to inquire as to why some fields are mostly or almost always blank.
Figure~\ref{fig:blank_fields} presents a graphic 
depiction of total empty (blank \& N/As) for each fields illustrating
the grouping into the Most, Partial, and Few categories.

@@

\begin{table}[tbp]
  \centering
  \caption{Negative duration of the largest magnitude ($-$730 days) excluding
    extreme negative values}
  \begin{tabular}{l l l r l}
    \toprule
    {created\_date} & {closed\_date} & {duration} 
    & \textbf{agency} \\
    \midrule
    2023-01-27 14:40:00 & 2022-01-14 14:40:00 & $-$378.0 & DOT \\
    2023-01-18 10:06:00 & 2022-01-12 10:06:00 & $-$371.0 & DOT \\
    2023-01-27 14:36:00 & 2022-01-22 14:35:00 & $-$370.0 & DOT \\
    2023-01-11 11:10:00 & 2022-01-09 11:10:00 & $-$367.0 & DOT \\
    2023-12-18 03:13:00 & 2023-01-16 13:10:00 & $-$335.6 & DOT \\
    \bottomrule
  \end{tabular}
  \label{tab:largest-errors}
\end{table}

@@

\textbf{landmark \& street\_name:} While the \texttt{landmark} field is 
intended for noteworthy locations (e.g., parks, airports), most entries 
are actually street names, with a 62\% match to \texttt{street\_name}. 
Even non-matches (excluding blanks) appear to be similar, differing 
only in spelling or formatting (e.g., ``NINTH AVE'' vs. ``9 AVE''). This 
indicates a possible misuse of the \texttt{landmark} field.

\textbf{cross\_street\_1 \& intersection\_street\_1 and cross\_street\_2 
\& intersection\_street\_2:} These two sets of street pairs are used to 
help identify the incident address. After applying address 
standardization using the R \textit{campfin} package, we found 88\% of 
these pairs to be duplicates, while 12\% do not match. This discrepancy 
raises the question of which field should be trusted. It is possible 
that the two pairs originated from different agencies that used 
different variable names to represent the same information, but no 
documentation exists to clarify this relationship.

\paragraph{Near Match}

\jy{This investigation could be relegated to on online supplement
  given the low percentage of near-match. It's taking too much spaces
  here.
}

\begin{table}[tbp]
    \centering
     \caption{Examples of near-matches for cross\_street\_1 and intersection\_street\_1.}
     \label{tab:x1nearmatches}
		\begin{tabular}{l l l r}
	        \toprule
	        \textbf{cross\_street\_1} & \textbf{intersection\_street\_1} 
	        & \textbf{agency} & \textbf{hamming\_distance} \\
	        \midrule
	        WEST 168 ST    & WEST 167 ST           & DOT    & 1 \\
	        105 AVE        & 107 AVE               & DOT    & 1 \\
	        145 ST         & 150 ST                & DOT    & 2 \\
	        138 ST         & 139 ST                & DOT    & 1 \\
	        19 AVE         & 20 AVE                & DOT    & 2 \\
	        76 ST          & 79 ST                 & DOT    & 1 \\
	        67 ST          & 68 ST                 & DOT    & 1 \\
	        \bottomrule
	    \end{tabular}
\end{table}

@@

In addition to the matching values, we encountered a number of 
\textit{near-matches}, situations where the \texttt{cross\_street} 
and \texttt{intersection\_street} are nearly, but not exactly, identical. 
An example of such a near-match is when one field reads ``HARMON DR'' 
and the other reads ``HARMON RD.'' To identify these near-matches, 
we applied the \href{https://en.wikipedia.org/wiki/Hamming_distance}
{\textit{Hamming Distance}}, which measures the number of character
\jy{needs a quick definition and reference}
changes required to make the two fields identical. For this analysis, 
we set the threshold for a near-match to a Hamming Distance of two (2).

Table~\ref{tab:x1near-matches} provides examples of near-matches for 
\texttt{cross\_street\_1} and \texttt{intersection\_street\_1}. These 
entries show slight variations in the street name or number, such as 
``WEST 168 ST'' and ``WEST 167 ST,'' both processed by DOT, where only 
one character differs. Although these discrepancies may seem small, 
they can have implications for data accuracy, particularly when 
geocoding or matching locations for analysis purposes.

Table~\ref{tab:summary1} presents a summary of the match and 
near-match results for \texttt{cross\_street\_1} and 
\texttt{intersection\_street\_1}. As the table shows, 88.15\% of the 
entries are exact matches, while 25.4\% are blank but matching. 
However, 11.85\% of the entries are non-matching, and a small fraction 
(0.002\%) fall into the near-match category. Although this percentage 
seems negligible, even a small number of near-matches can lead to 
confusion or errors when analyzing location-based data, especially 
if critical decisions rely on accurate street data. Address 
standardization or correction processes may be necessary to resolve 
these inconsistencies and improve the reliability of the data.

>>>>

The Mostly Empty category includes such fields as
\texttt{taxi\_company\_borough}, \texttt{due\_date},
\texttt{road\_ramp}, and \texttt{bridge\_highway\_name}.
The Partially Empty includes such fields as
\texttt{location\_type}, \texttt{landmark}, 
and \texttt{cross\_street\_1 \& \_2}. And the Few/None Empty includes
\texttt{created\_date}, \texttt{complaint\_type},
\texttt{agency}, and \texttt{status}. In some analysis efforts, it may be
prudent to inquire as to why some fields are mostly or almost always blank.
Figure~\ref{fig:blank_fields} presents a graphic 
depiction of total empty (blank \& N/As) for each fields illustrating
the grouping into the Most, Partial, and Few categories.

@@

\begin{table}[tbp]
  \centering
  \caption{Sample  negative durations [excluding extreme  values ($-$730 days)]}
  \begin{tabular}{l l l r l}
    \toprule
    {created\_date} & {closed\_date} & {duration} 
    & \textbf{agency} \\
    \midrule
    2023-01-27 14:40:00 & 2022-01-14 14:40:00 & $-$378.0 & DOT \\
    2023-01-18 10:06:00 & 2022-01-12 10:06:00 & $-$371.0 & DOT \\
    2023-01-27 14:36:00 & 2022-01-22 14:35:00 & $-$370.0 & DOT \\
    2023-01-11 11:10:00 & 2022-01-09 11:10:00 & $-$367.0 & DOT \\
    2023-12-18 03:13:00 & 2023-01-16 13:10:00 & $-$335.6 & DOT \\
    \bottomrule
  \end{tabular}
  \label{tab:largest-errors}
\end{table}

@@

\texttt{landmark} \& \texttt{street\_name:} While the \texttt{landmark} field is 
intended for noteworthy locations (e.g., parks, airports), most entries 
are actually street names, with a 62\% match to \texttt{street\_name}. 
Even non-matches (excluding blanks) appear to be similar, differing 
only in spelling or formatting (e.g., ``NINTH AVE'' vs. ``9 AVE''). This 
indicates a possible misuse of the \texttt{landmark} field as well as
a high level of redundancy.

\texttt{cross\_street\_1} \& \texttt{intersection\_street\_1} and \texttt{cross\_street\_2} 
\& \texttt{intersection\_street\_2}: These two sets of street pairs are used to 
help identify the incident address. We found 88\% of 
these pairs to be duplicates. This duplication  
raises the question of which field-pair should be trusted. It is possible 
that the two street pairs originated from different agencies that used 
different variable names to represent the same information, but no 
documentation exists to clarify this relationship. Without question
these two street-pairs are highly redundant.

@@

Table~\ref{tab:summary1} presents a summary of the match and 
near-match results for \texttt{cross\_street\_1} and 
\texttt{intersection\_street\_1}. As the table shows, 88.15\% of the 
entries are exact matches, while 25.4\% are blank but matching.
